---
title: "Ranking Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(weightsetdecomp)
library(gridExtra)
```

## Examples from Publication Perini 2025


### Figure 1

This introduces the basic visualizations created by this library: a weight set decomposition and its dual graph when aggregating multiple ranked lists using a weighted method. The dual graph visualization (right) is a simplification of the more complete weight set decomposition (left); however, it is generalizable to more than three input ranks/scores.

```{r}
Lambda <- weight_set()
rankdf <- data.frame('risk1'=c(1,2,3,4,5), 
           'risk2'=c(2,3,1,5,4), 'risk3'=c(3,1,5,4,2))
Lambda <- rank_aggregation_grid(Lambda=Lambda,metrics=rankdf)
outlist <- rank_decomposition_exact(input_data=rankdf,Lambda)
g1 <- plot_aggregation_exact(Lambda,outlist,
          with_lines=TRUE,with_points=FALSE,plotly_text=FALSE,leg.pos='left',random.color=0)
g2 <- plot_ggnetwork_grid(Lambda, metrics=rankdf, node_size=20, node_label=TRUE,
                    edge_size=1.5,random.color=0)
grid.arrange(g1,g2,ncol=2)
```



### Example 1.1

By default, the flex_rank() function will break ties lexicographically, i.e.,  by order of index. The appropriate way to interpret the first output vector $r=[1, 3, 5, 4, 2]$ is the following: 

* $r_1=1$ indicates that the first investment is ranked first; 
* $r_2=3$ indicates that the second investment is in third place; 
* $r_3=5$ indicates that the third investment is in last place; 
* $r_4=4$ indicates that the fourth investment is in fourth place; and 
* $r_5=2$ indicates that the final investment is in second place.

In the weightsetdecomp library, this labeling will be referred to as _Rank.Label_. 

```{r}
risk <- c(1.0, 2.0, 3.1, 2.0, 1.7)
flex_rank(risk)
flex_rank(risk,ties=TRUE)
```

### Example 1.4

The following functions compute the line segments between all pairs of investments and plots one desired line segment. For one incomparable pair $i,j\in[n]$, the line segment represents the intersection $H(i,j)\cap \Lambda$, which is described by its two endpoints.
Computing all line segments results in the complete weight set decomposition. 

```{r}
Lambda <- weight_set()
rankdf <- data.frame('risk1'=c(1,2,3,4,5), 
           'risk2'=c(2,3,1,5,4), 'risk3'=c(3,1,5,4,2))
Lambda <- rank_aggregation_grid(Lambda=Lambda,metrics=rankdf)
outlist <- rank_decomposition_exact(metrics,Lambda)
Linedf <- outlist[[1]]
# First, compare investments 1 and 2
plot_lineseg_grid(Lambda,Linedf,item_pair=c(1,2),
    annotations=TRUE)
# Second, compare investments 1 and 3
plot_lineseg_grid(Lambda,Linedf,item_pair=c(1,3),
    annotations=TRUE)
# Finally, plot all line segments for the complete decomposition
plot_aggregation_exact(Lambda,outlist,
    with_lines=TRUE,with_points=TRUE,leg.pos='bottom')
```

### Example 1.5

A tie is induced in the first scoring vector (``risk1'' column of rankdf). This results in the line segment having the corresponding vertex as an endpoint. 

```{r}
Lambda <- weight_set()
rankdf <- data.frame('risk1'=c(1,1,3,4,5), 
           'risk2'=c(2,3,1,5,4), 'risk3'=c(3,1,5,4,2))
Lambda <- rank_aggregation_grid(Lambda=Lambda,metrics=rankdf)
outlist <- rank_decomposition_exact(metrics,Lambda)
Linedf <- outlist[[1]]
# First, compare investments 1 and 2 (now with a tie)
plot_lineseg_grid(Lambda,Linedf,item_pair=c(1,2),
    annotations=TRUE)
# Finally, plot all line segments for the complete decomposition
plot_aggregation_exact(Lambda,outlist,
    with_lines=TRUE,with_points=TRUE,leg.pos='bottom')
```

### Example 3.3

When focusing on a single investment, the following functions plot the investment's ranked position for all weighted aggregations. 
First, for a small portfolio of five investments, plot the ranked position of  investment 1. 
Second, for a large portfolio of 1,000 investments, plot exactly where investment 1 is ranked in 434th position. 

```{r}
# First, for small portfolio
Lambda <- weight_set()
rankdf <- data.frame('risk1'=c(1,2,3,4,5), 
           'risk2'=c(2,3,1,5,4), 'risk3'=c(3,1,5,4,2))
Lambda <- rank_aggregation_grid(Lambda=Lambda,metrics=rankdf)
plot_position_grid(Lambda, item=1,leg.pos='bottom')
# Second, for large portfolio
data("sampledf")
Lambda <- rank_aggregation_grid(Lambda=Lambda,metrics=sampledf)
plot_position_grid(Lambda, item=1, position=434)
```

### Example 3.6

Rather than distinguishing every indifference region by its full ranked order, identify only its top-$\alpha$-ranked investments. This is useful for large portfolios, as in this example with 1,000 investments.

```{r}
Lambda <- weight_set()
data("sampledf")
# First, for top-1 regions
Lambda <- rank_topalpha_grid(Lambda,alpha=1,sampledf)
plot_aggregation_grid(Lambda,by='TopAlpha',leg.pos = 'bottom')
# Second, for top-10 regions
Lambda <- rank_topalpha_grid(Lambda,alpha=10,sampledf)
plot_aggregation_grid(Lambda,by='TopAlpha')
```

### Example 3.9

It is possible to apply a simple filtering of a budget constraint where each investment has a given cost. For each indifference region, the investments are considered in ranked order and are ``purchased'' if it fits within the current budget. This represents a greedy approximate solution to the knapsack problem, so it may be suboptimal compared to the true knapsack solution. 

```{r}
Lambda <- weight_set(stepsize=0.01)
rankdf <- data.frame('risk1'=c(1,2,3,4,5), 
           'risk2'=c(2,3,1,5,4), 'risk3'=c(3,1,5,4,2))
costs <-c(5.9, 4.4, 2.2, 5.5, 3.6) #sample(seq(2,6,0.1),5)
Lambda <- rank_budget_grid(Lambda, costs, budget=10, metrics=rankdf)
plot_aggregation_grid(Lambda,by='Budget',leg.pos='bottom')
```




